---
title: "knitr-tools Demo"
output:
  pdf_document: default
  html_document: default
---

## Code

### Chunk knitting timing

Print timing information for each chunk. This also collects and stores the
knitting times printed in the final report.

_This is the core of the knitr timing code. It is required by the other pieces._

It contains the following customizable settings:

Setting | Function
--- | ---
CHUNK_TIME_CUTOFF | exclude chunks with knitting times longer than this (in minutes)
knitr::opts_chunk$set(time_it= ) | apply the timing hook to all chunks (TRUE/FALSE)

**TODO: suppress per-chunk printing but still print report**

```{r "knitr chunk timing"}
# Print run times for all chunks

# Enable timing for all chunks
knitr::opts_chunk$set(time_it=TRUE)

# Skip chunks faster than this in final report
CHUNK_TIME_CUTOFF <- 0.01

chunk_times <- list()

# Custom timing chunk hook
knitr::knit_hooks$set(time_it = local({
  start_time <- NULL
  chunk <- NULL

  function(before, options) {
    if (before) {
      # record the current time before the chunk
      start_time <<- Sys.time()
    } else {
      # calculate the time difference after the chunk
      chunk <- options$label
      chunk_time <- difftime(Sys.time(), start_time, units = "mins")

      chunk_times[chunk] <<- chunk_time
      print(sprintf("Chunk '%s' took %.2f min", chunk, chunk_time))
    }
  }
}))
```

### Knitting timing report

Print a report when knitting is complete that shows how long all
chunks took to knit. This is useful to quickly identify which chunks
are slowing things down.

_This block can be removed if you don't want to see the report at the end._

It contains the following customizable settings:

Setting | Function
--- | ---
CHUNK_NAME_WIDTH | chunk name padding for final report to align times (negative integer)

```{r "print knitr timing report"}

# Hook to print timing report
local({
  # Increase this to align the output for longer chunk names
  CHUNK_NAME_WIDTH=-50
  
  # Preserve existing hook
  old_document_hook <- knitr::knit_hooks$get("document")

  knitr::knit_hooks$set(document = function(x) {
    print("Chunk knitting time report")
    print(sprintf("(longer than %.2f min)", CHUNK_TIME_CUTOFF))
    cat("\n")

    for (chunk in names(chunk_times)) {
      if (chunk_times[[chunk]] > CHUNK_TIME_CUTOFF) {
        print(sprintf("Chunk %*s : %5.2f min", CHUNK_NAME_WIDTH, chunk, chunk_times[[chunk]]))
      }
    }
  cat("\n")

  # Run any document hook that was already defined
  old_document_hook(x)
  })
})
```

### Knitting completion timestamp

Print a timestamp when knitting is complete. 

_This block can be removed if you don't want to see the timestamp at the end._

```{r "print knitr completion timestamp"}
# Print the time after knitting is complete
local({
  # Preserve existing hook
  old_document_hook <- knitr::knit_hooks$get("document")

  knitr::knit_hooks$set(document = function(x) {
  cat("\n")
  print(paste("Knitting completed at", Sys.time()))
  cat("\n")

  # Run any document hook that was already defined
  old_document_hook(x)
  })
})
```
**TODO: Add total runtime**

# Demonstration of the chunk timing output
### Example output

The following output will appear in the **Rstudio** **R Markdown** window

#### A 3-second dummy chunk

```{r "Dummy Chunk #1 - 3 seconds", echo=FALSE}
Sys.sleep(3)
```


```
  |...............................................                       |  67%
label: Dummy Chunk #1 - 3 seconds
[1] "Chunk 'Dummy Chunk #1 - 3 seconds' took 0.05 min"
```

#### A chunk that knits instantly

```{r "Dummy Chunk #2 - 0 seconds", echo=FALSE}
invisible()
```

Output:

```
  |........................................................              |  80%
label: Dummy Chunk #2 - 0 seconds
[1] "Chunk 'Dummy Chunk #2 - 0 seconds' took 0.00 min"
```
#### A longer-running chunk

```{r "Dummy Chunk #3 - random seconds", echo=FALSE}
Sys.sleep(runif(1,1,10))
```

Output:

```
  |......................................................................| 100%
label: Dummy Chunk #3 - random seconds
[1] "Chunk 'Dummy Chunk #3 - random seconds' took 0.15 min"
```

### The knitting time report

When knitting is complete a report is printed. Only chunks which took longer than
`r CHUNK_TIME_CUTOFF` minutes will be printed.

```
[1] "Chunk run time report"
[1] "(longer than 0 min)"

[1] "Chunk Dummy Chunk #1 - 3 seconds     :  0.05 min"
[1] "Chunk Dummy Chunk #2 - 0 seconds     :  0.00 min"
[1] "Chunk Dummy Chunk #3 - random seconds:  0.15 min"
```
A timestamp will also be printed upon completion. This is useful if you can't
remember how long ago you last knitted your document.

```
[1] "Knitting completed at 2020-07-31 16:26:16"
```